---
title: "scRNA data analysis"
output: html_notebook
---

Reading in the data
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
# Here 'ie' stands for intestinal epithelium
ie1.data <- Read10X(data.dir = "~/Desktop/study/R/ie/GSM2836573/")
ie2.data <- Read10X(data.dir = "~/Desktop/study/R/ie/GSM2836574/")
ie3.data <- Read10X(data.dir = "~/Desktop/study/R/ie/GSM2836575/")
ie4.data <- Read10X(data.dir = "~/Desktop/study/R/ie/GSM2836576/")
ie5.data <- Read10X(data.dir = "~/Desktop/study/R/ie/GSM2836577/")
ie6.data <- Read10X(data.dir = "~/Desktop/study/R/ie/GSM2836578/")
ie1 <- CreateSeuratObject(counts = ie1.data, project = "Duo_M1", min.cells = 20, min.features = 0)
ie2 <- CreateSeuratObject(counts = ie2.data, project = "Duo_M2", min.cells = 20, min.features = 0)
ie3 <- CreateSeuratObject(counts = ie3.data, project = "Jej_M1", min.cells = 20, min.features = 0)
ie4 <- CreateSeuratObject(counts = ie4.data, project = "Jej_M2", min.cells = 20, min.features = 0)
ie5 <- CreateSeuratObject(counts = ie5.data, project = "Il_M1", min.cells = 20, min.features = 0)
ie6 <- CreateSeuratObject(counts = ie6.data, project = "Il_M2", min.cells = 20, min.features = 0)
ie <- merge(ie1, y = c(ie2, ie3, ie4, ie5, ie6), add.cell.ids = c("Duo_M1", "Duo_M2", "Jej_M1", "Jej_M2", "Il_M1", "Il_M2"), project = 'all')
ie
head(colnames(ie))
```

```{r}
table(ie$orig.ident)
ie[["percent.mt"]] <- PercentageFeatureSet(ie, pattern = "^mm10-mt-")
VlnPlot(ie, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.5)
```

```{r}
plot1 <- FeatureScatter(ie, feature1 = "nCount_RNA", feature2 = "percent.mt") + theme(axis.text.x = element_text(angle = 90))
plot2 <- FeatureScatter(ie, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + theme(axis.text.x = element_text(angle = 90))
plot1 + plot2
```

Filtering (using the same parameters in scRNA_practice.ipynb)
```{r}
ie <- subset(ie, subset = nFeature_RNA >= 700 & nCount_RNA >= 1500 & nCount_RNA <= 40000 & percent.mt < 20)
```


Dimensional reduction without batch correction
```{r}
ie <- NormalizeData(ie, normalization.method = "LogNormalize", scale.factor = 10000)
ie <- FindVariableFeatures(ie, selection.method = "vst", nfeatures = 2000)
ie <- ScaleData(object = ie, verbose = FALSE)
ie <- RunPCA(object = ie, npcs = 50, verbose = FALSE)
ie <- RunUMAP(object = ie, reduction = "pca", dims = 1:50)
p1 <- DimPlot(object = ie, reduction = "umap", group.by = "orig.ident")
plot_grid(p1)
```
As can be seen, the batches are not well mixed.

Batch correction
```{r}
ie.list <- SplitObject(object = ie, split.by = "orig.ident")
for (i in 1:length(x = ie.list)) {
    ie.list[[i]] <- NormalizeData(object = ie.list[[i]], verbose = FALSE)
    ie.list[[i]] <- FindVariableFeatures(object = ie.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
reference.list <- ie.list[c("Duo_M1", "Duo_M2", "Jej_M1", "Jej_M2", "Il_M1", "Il_M2")]
ie.anchors <- FindIntegrationAnchors(object.list = reference.list, dims = 1:50)
ie.integrated <- IntegrateData(anchorset = ie.anchors, dims = 1:50)
DefaultAssay(object = ie.integrated) <- "integrated"
```

Dimensional reduction and visualization
```{r}
ie.integrated <- ScaleData(object = ie.integrated, verbose = FALSE)
ie.integrated <- RunPCA(object = ie.integrated, npcs = 50, verbose = FALSE)
ie.integrated <- RunUMAP(object = ie.integrated, reduction = "pca", dims = 1:50)
p2 <- DimPlot(object = ie.integrated, reduction = "umap", group.by = "orig.ident")

plot_grid(p2)
```
All batches are well mixed, maybe it is not very good since it is possible that biological differences among Duo, Jej and Il are also removed. In scRNA_practice.ipynb (combat is used there for batch correction), for example, Duo, Jej belong to Enterocyte (Proximal) and Il belongs to Enterocyte (Distal).

Clustering. Here resolution is chosen as 0.2 which results in the same number of clusters as that in scRNA_practice.ipynb. But it leads to different annotations as we can see below.
```{r}
ie.integrated <- FindNeighbors(ie.integrated, dims = 1:10)
ie.integrated <- FindClusters(ie.integrated, resolution = 0.2)
DimPlot(ie.integrated, reduction = "umap")
```

Find marker genes
```{r}
ie.integrated.markers <- FindAllMarkers(ie.integrated, only.pos = TRUE, min.pct = 0.2, logfc.threshold = 0.2)
```
```{r}
ie.integrated.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
```

Check early enterocyte marker Arg2
```{r}
VlnPlot(ie.integrated, features = c("mm10-Arg2"))
```
It looks both clusters 4 and 6 have high expression of Arg2. 

```{r}
features.plot <- c("mm10-Arg2")
FeaturePlot(object = ie.integrated, features = features.plot)
```

Check some other enterocyte markers.
```{r}
FeaturePlot(ie.integrated, features = c("mm10-Alpi", "mm10-Apoa1", "mm10-Apoa4", "mm10-Fabp1"))
```
Compared with markers in scRNA_practice.ipynb, clusters 4 and 6 should be Enterocyte imm. and Enterocyte mature respectively.

```{r}
VlnPlot(ie.integrated, features = "percent.mt", pt.size=0)
FeaturePlot(ie.integrated, features = "percent.mt")
```
No clusters have obvious higher mitochondrial read fractions from violin plot. From feature plot, cluster 0 has higher mitochondrial read fractions than other clusters. Cluster 0 might be EP.

Check Stem markers.
```{r}
FeaturePlot(ie.integrated, features = c("mm10-Gkn3", "mm10-Ascl2", "mm10-Lgr5", "mm10-Olfm4", "mm10-Axin2", "mm10-Slc12a2"))
```
Cluster 2 is Stem.

Compared with marker genes in scRNA_practice.ipynb, cluster 1 is TA, cluster 3 is Goblet, cluster 5 is Paneth, cluster 7 is Enteroendocrine, cluster 8 is Tuft.

Annotate clusters
```{r}
new.cluster.ids <- c("EP", "TA", "Stem", "Goblet", "Enterocyte imm.", "Paneth", "Enterocyte mature", "Enteroendocrine", "Tuft")
names(new.cluster.ids) <- levels(ie.integrated)
ie.integrated <- RenameIdents(ie.integrated, new.cluster.ids)
DimPlot(ie.integrated, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```
Here, different from scRNA_practice.ipynb, after the first round of clustering, EP belongs to one cluster while Enterocyte have two clusters already.
